---
title: "Cluster Analysis"
output:
  word_document: default
  html_document: default
---

## Cluster Analysis
# create 2 arrays and convert to data frame with 2 obervations

```{r two_players, echo = FALSE}

player_list = c('player1','player2')
x = c(5,15)
y = c(4,10)
two_players = data.frame(x, y, row.names = player_list)
head(two_players)
```


# Plot the positions of the players

```{r two_players plot, echo=FALSE}

library(ggplot2)
ggplot(two_players, aes(x = x, y = y)) + 
  geom_point() +
  # Assuming a 40x60 field
  lims(x = c(-30,30), y = c(-20, 20))

#add coordinate labels on the plot
```


#Calculate Euclidean distance between players
```{r distance, echo=FALSE}

# Split the players data frame into two observations
player1 <- two_players[1, ]
player2 <- two_players[2, ]
player1
player2



# Calculate and print the distance using the Euclidean Distance formula
player_distance <- sqrt( (player1$x - player2$x)^2 + (player1$y - player2$y)^2 )
player_distance


###Use Distance fucntion to create matrix
# Calculate the Distance Between two_players
dist_two_players <- dist(two_players)
dist_two_players


```


#Add a 3rd player
```{r 3rd player,echo=FALSE}


player_list = c('player1','player2','player3')
x = c(5,15,0)
y = c(4,10,20)
three_players = data.frame(x, y, row.names = player_list)

ggplot(three_players, aes(x = x, y = y)) + 
  geom_point() +
  # Assuming a 40x60 field
  lims(x = c(-30,30), y = c(-20, 20))

# Calculate the Distance Between three_players
dist_three_players <- dist(three_players)
dist_three_players

```



#Use cluster analysis to find teams at opening kickoff
#create FULL lineups in a 6v6 game 
```{r full game, echo=FALSE}


x = c(-1,-2,8,7,-12,-15,-13,15,21,12,-25,26)
y = c(1,-3,6,-8,8,0,-10,16,2,-15,1,0)
lineup = data.frame(x,y)
head(lineup)
tail(lineup)

#View Field
ggplot(lineup, aes(x = x, y = y),) + 
  geom_point() +
  # Assuming a 40x60 field
  lims(x = c(-30,30), y = c(-20, 20))


# Calculate the Distance
dist_players = dist(lineup)


```


#Use cluster analysis to identify players of each team (current position is just prior to opening kickoff)

```{r application, echo=FALSE}
# Perform the hierarchical clustering using the complete linkage
hc_players = hclust(dist_players, method = "complete")


# Calculate the assignment vector with a k of 2

clusters_k2 = cutree(hc_players, k = 2)

# Create a new data frame storing these results
library(dplyr)
lineup_k2_complete = mutate(lineup, cluster = clusters_k2)
lineup_k2_complete
validate = lineup_k2_complete %>% group_by(cluster) %>% tally()
validate

#solution?
ggplot(lineup_k2_complete, aes(x = x, y = y),) + 
  geom_point(aes(colour = factor(cluster))) +
  # Assuming a 40x60 field
  lims(x = c(-30,30), y = c(-20, 20))

```

#Dendrogram Deep Dive
```{r dendrogram, echo=FALSE}
library(dendextend)
#Vizualize the process
#Dendrogram
plot(hc_players)

#Compare methods: complete, single, average
hc_complete <- hclust(dist_players, method = "complete")
hc_single <- hclust(dist_players, method = "single")
hc_average <- hclust(dist_players, method = "average")

# Plot & Label the 3 Dendrograms Side-by-Side
par(mfrow = c(1,3))
plot(hc_complete, main = 'Complete Linkage')
plot(hc_single, main = 'Single Linkage')
plot(hc_average, main = 'Average Linkage')
```

#Dendrogram Deep Dive
```{r deep dive 2, echo=FALSE}

dist_players <- dist(lineup, method = 'euclidean')
hc_players <- hclust(dist_players, method = "complete")

# Create a dendrogram object from the hclust variable
dend_players <- as.dendrogram(hc_players)

# Plot the dendrogram
plot(dend_players)


# Color branches by cluster formed from the cut at a height of 20 & plot
dend_20 <- color_branches(dend_players, h = 20)

# Plot the dendrogram with clusters colored below height 20
plot(dend_20)

# Color branches by cluster formed from the cut at a height of 40 & plot
dend_40 <- color_branches(dend_players, h = 40)

# Plot the dendrogram with clusters colored below height 40
plot(dend_40)

#Slice clusters by dendrogram height

dist_players <- dist(lineup, method = 'euclidean')
hc_players <- hclust(dist_players, method = "complete")

# Calculate the assignment vector with a h of 20
clusters_h20 <- cutree(hc_players, h = 20)

# Create a new data frame storing these results
lineup_h20_complete <- mutate(lineup, cluster = clusters_h20)

# Calculate the assignment vector with a h of 40
clusters_h40 <- cutree(hc_players, h = 40)

# Create a new data frame storing these results
lineup_h40_complete <- mutate(lineup, cluster = clusters_h40)

# Plot the positions of the players and color them using their cluster for height = 20
ggplot(lineup_h20_complete, aes(x = x, y = y, color = factor(cluster))) +
  geom_point()

# Plot the positions of the players and color them using their cluster for height = 40
ggplot(lineup_h40_complete, aes(x = x, y = y, color = factor(cluster))) +
  geom_point()

```

#Kmeans Clustering

```{r kmeans, echo=FALSE}
# Build a kmeans model from lineup data frame
model_km2 <- kmeans(lineup, centers = 2)

# Extract the cluster assignment vector from the kmeans model
clust_km2 <- model_km2$cluster

# Create a new data frame appending the cluster assignment
lineup_km2 <- mutate(lineup, cluster = clust_km2)
# Plot the positions of the players and color them using their cluster
ggplot(lineup_km2, aes(x = x, y = y, color = factor(cluster))) +
  geom_point()

```


##More Kmeans
```{r more kmeans, echo=FALSE}
library(purrr)

# Use map_dbl to run many models with varying value of k (centers)
tot_withinss <- map_dbl(1:10,  function(k){
  model <- kmeans(x = lineup, centers = k)
  model$tot.withinss
})

# Generate a data frame containing both k and tot_withinss
elbow_df <- data.frame(
  k = 1:10,
  tot_withinss = tot_withinss
)
# Plot the elbow plot
ggplot(elbow_df, aes(x = k, y = tot_withinss)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = 1:10)

```

#Kmeans with silhouette analysis

```{r silhouette, echo=FALSE}

library(cluster)

# Generate a k-means model using the pam() function with a k = 2
pam_k2 <- pam(lineup, k = 2)

# Plot the silhouette visual for the pam_k2 model
plot(silhouette(pam_k2))

# Generate a k-means model using the pam() function with a k = 3
pam_k3 <- pam(lineup, k = 3)

# Plot the silhouette visual for the pam_k3 model
plot(silhouette(pam_k3))


```


#Scaling Features/Inputs for Cluster Analysis 
```{r factoral, echo=FALSE}

#Import trees data set
data("trees")
head(trees)
#subset 3 trees
three_trees = trees[c(1:3),]

#unscaled distance
dist_trees = dist(three_trees)
#scaled distance
scaled_three_trees = scale(three_trees)
dist_scaled_trees = dist(scaled_three_trees)


# Output the results of both Matrices
print('Without Scaling')
dist_trees

print('With Scaling')
dist_scaled_trees

```

#Dealing with factoral values
library(dummies)

#create survey data frame
job_satisfaction = c('low','low','high','low','mid')
is_happy = c('no','no','yes','no','no')
job_survey = data.frame(job_satisfaction, is_happy)
summary(job_survey)
head(job_survey)


# Dummify the Survey Data
dummy_survey <- dummy.data.frame(job_survey)
print(job_survey)
print(dummy_survey)

# Calculate the Distance
dist_survey <- dist(dummy_survey, method = 'binary')

# Print the Distance Matrix
print(dist_survey)



